---
title: "lidarSHM"
author: "Aland Chan"
date: "2025-06-04"
output: github_document
---

## lidarSHM

This package is designed for calculating various lidar metrics from lidar point clouds and canopy height models.

The package has three main functions:

1. `lidar_SHM` (in development) calculates a shrub height model (SHM) from lidar point cloud data.

2. `CHM_metrics` calculates a number of useful structural metrics from a canopy height model or shrub height model in one single step. The function takes a CHM SpatRaster and a vector of metric names; returns a SpatRaster with multiple layers (names of layers corresponding to metric names). Definitions of metrics are based on Zhang et al. (2024) with several minor modifications and additions.

3. `liDAR_metrics` calculates a number of useful structural metrics from point clouds in one single step. The function takes either a LAS or LAScatalog object created with the `lidR::` package as inputs. If a `LAS` object is supplied, the function returns a multi-layered SpatRaster, with layers corresponding to metrics users requested. Alternatively, if a `LAScatalog` object is supplied, it writes the SpatRasters as `.tif` files in folder specified by `opt_output_files()`. By default, the function performs mcc ground classification, but users could also supply a ground-classified point cloud if they want to skip this time-consuming step or prefer another ground classification algorithm. For the function to operate properly, las@data should have the columns named `Z` (height), `ReturnNumber`, `NumberOfReturns`, and `ScanAngleRank` (angle in degrees). The last two are specifically used for ePAI calculations. The definitions and names of metrics are mostly based on Shokirov et al. (2023). 

## List of metrics

Some of the groupings and definitions are derived from Zhang et al. (2024) and Shokirvo et al. (2023).

```{r,echo = F}

library(knitr)

metric_list<-read.csv("F:/Cairngorms_project/LiDAR/LiDAR_metrics/250306_LiDAR_metrics_list_package.csv")

metric_list$Function<-gsub("CHM","CHM_metrics",metric_list$Function)
metric_list$Function<-gsub("LiDAR","LiDAR_metrics",metric_list$Function)

kable(metric_list)

```


## Installation

```{r, eval = F}

library(devtools)

devtools::install_github("alandhyc/lidarSHM",force = T)

```

## Examples

### CHM_metrics

Calculating CHM metrics from a CHM.

```{r,eval=F}

library(terra)

#

chm_r<-terra::rast("path/to/chm")

#Calculating all metrics
#Aggregate 50x50 pixels into one (resotuion 1m for CHM, 50m for metrics)
#Defining "gap" as things under 2m

chm_metrics<-CHM_metrics(chm = chm_r,
                         agg = 50,
                         gap_thres = 2,
                         metrics = "all")

#Calculating specific metrics
#Just do max height and standard deviation

chm_metrics_subset<-CHM_metrics(chm = chm_r,
                                agg = 50,
                                gap_thres = 2,
                                metrics = c("chm_maxH","chm_sdH"))

```

### LiDAR_metrics

Applying the function across a `LAS` object

```{r, eval = F}

library(lidR)

las<-readLAS("/path/to/las/file.las")

#With all options
#Perform ground classification
#Resolution at 50m
#Calculate most metrics based on points >1.3m ("canopy" definition)
#Anything <2m treated as a gap
#ground classification with mcc using mcc_s as s and mcc_t as t
#Maximum height of tree approximately 35m
#height classes -1-2, 2-5, 5-10, 10-15, 15-35 when calculating shannon index
#voxel resolution 0.5 m3 when calculating canopy volumn 
#Stratifying point cloud into three layers (0-1, 1-10, 10-35) when calculated layer-specific metrics in Shakirov et al. (2023)
#Calculate all metrics

output_metrics<-LiDAR_metrics(las,
                              ground_classified = F,
                              res=50,
                              h_cutoff=1.3,
                              gap_thres = 2,
                              mcc_s=1.5,
                              mcc_t=0.3,
                              zmax = 35,
                              shannon_cut=c(-1,2,5,10,15,35),
                              vox_res=0.5,
                              L1_range=c(0,1),
                              L2_range=c(1,10),
                              L3_range=c(10,35),
                              metrics = "all"
                              )

#Calculating specific metrics
#Point cloud already ground classified

output_metrics<-LiDAR_metrics(las,
                              ground_classified = T,
                              res=50,
                              zmax = 35,
                              metrics = c("lidar_maxH","lidar_ePAI")
                              )


```

Alternatively, we could apply the function to a `LAScatalog` object with parallel computing and output written to a folder. 

```{r, eval = F}

library(lidR)
library(lidRmetrics)
library(future)

#Set up parallel computation
ncores<-5
plan(multisession, workers = ncores)
set_lidr_threads(ncores)

#Load las catalog

ctg<-readLAScatalog("/path/to/las/files")

output_dir<-"/path/to/output/directory"


opt_output_files(ctg)<-paste0(output_dir,"/{ORIGINALFILENAME}")
opt_chunk_buffer(ctg)<-20 #Set a buffer

LiDAR_metrics(las,
              ground_classified = F,
              res=50,
              h_cutoff=1.3,
              gap_thres = 2,
              mcc_s=1.5,
              mcc_t=0.3,
              zmax = 35,
              shannon_cut=c(-1,2,5,10,15,35),
              vox_res=0.5,
              L1_range=c(0,1),
              L2_range=c(1,10),
              L3_range=c(10,35),
              metrics = "all"
              )

```


## Acknowledgements

Evie Huhtala for providing part of the code in the `CHM_metrics()` function

## References
Zhang, B., Fischer, F.J., Prober, S.M., Yeoh, P.B., Gosper, C.R., Zdunic, K. and Jucker, T. (2024), Robust retrieval of forest canopy structural attributes using multi-platform airborne LiDAR. Remote Sens Ecol Conserv, 10: 725-742. https://doi.org/10.1002/rse2.398

Shokirov, S. et al. (2023) ‘Habitat highs and lows: Using terrestrial and UAV LiDAR for modelling avian species richness and abundance in a restored woodland’, Remote Sensing of Environment. Elsevier, 285, p. 113326. doi: 10.1016/J.RSE.2022.113326.

## Contact

For questions or suggestions, open an issue or contact ```hyc43@cam.ac.uk```
